Analysis of epistatic inflation under large single additive effects, with a focus on the replication rate

### 2020-01-23 null_replication/docs/analyse_results.rmd
updated

load("../data/aggregate.rdata")
res$polygenic <- res$sim %% 2 == 0
res$gene <- sapply(res$i, function(x) strsplit(x, "_")[[1]][1])
res$scenario <- 2
l[[2]] <- res

load("../data/aggregate2.rdata")
res$polygenic <- res$sim %% 2 == 0
res$gene <- sapply(res$i, function(x) strsplit(x, "_")[[1]][1])
res$scenario <- 3
l[[3]] <- res

@@

repl_rate <- bind_rows(
	group_by(res, varexp_bin, scenario) %>% summarise(m=mean(bonf_bonf_rr, na.rm=T)) %>% mutate(threshold="Bonferroni"),
	group_by(res, varexp_bin, scenario) %>% summarise(m=mean(bonf_fdr_rr, na.rm=T)) %>% mutate(threshold="FDR"),
	group_by(res, varexp_bin, scenario) %>% summarise(m=mean(bonf_bonf2_rr, na.rm=T)) %>% mutate(threshold="Experiment")
)
repl_rate %>% ggplot(aes(x=varexp_bin, y=m, group=threshold)) +
geom_line(aes(colour=threshold)) +
scale_colour_brewer(type="qual") +
theme(axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5)) +
labs(x="Variance explained by cis additive variant", y="Replication rate") +
facet_grid(. ~ scenario) +
ylim(c(0,0.3))
```

@@

>>>>

load("../data/aggregate2.rdata")
res$polygenic <- res$sim %% 2 == 0
res$gene <- sapply(res$i, function(x) strsplit(x, "_")[[1]][1])
res$scenario <- 2
l[[2]] <- res

load("../data/aggregate3.rdata")
res$polygenic <- res$sim %% 2 == 0
res$gene <- sapply(res$i, function(x) strsplit(x, "_")[[1]][1])
res$scenario <- 3
l[[3]] <- res

@@

repl_rate <- bind_rows(
	group_by(res, varexp_bin, scenario) %>% summarise(m=mean(bonf_bonf_rr, na.rm=T)) %>% mutate(threshold="Bonferroni"),
	group_by(res, varexp_bin, scenario) %>% summarise(m=mean(bonf_fdr_rr, na.rm=T)) %>% mutate(threshold="FDR"),
	group_by(res, varexp_bin, scenario) %>% summarise(m=mean(bonf_bonf2_rr, na.rm=T)) %>% mutate(threshold="Experiment")
)
repl_rate %>% ggplot(aes(x=varexp_bin, y=m, group=threshold)) +
geom_line(aes(colour=threshold)) +
scale_colour_brewer(type="qual") +
theme(axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5)) +
labs(x="Variance explained by cis additive variant", y="Replication rate") +
facet_grid(. ~ scenario)
```

@@

## Replication rate of MBNL1

Analysing the MBNL1 locus - what is the discovery and replication rate?

There are 40000 simulations, where the cis-additive variant explains some amount of the variance in the phenotype (ranging from 0% to 50%). If a signal is found to be significant then it is tested for replication. In this plot, the x axis shows how many independent associations found to be significant for a simulations, and the y axis is the proportion of those that replicate at p < 0.05/501. The simulations are binned based on the simulated additive variance (columns of plots). The blue dots are from the simulations, and the red dot is the empirical result for MBNL1. Note that the cis-additive variant explains 10.5% of the variance of MBNL1, so the relevent boxes for comparison against the simulations are the second and third from the left. There are very few discovered loci here, and the replication rate of those is always zero. When the additive locus explains more variance, then the number of discovered hits increases and the replication rate increases.

```{r, fig.width=12}
temp <- subset(res, scenario == 1)

temp$vbin <- cut(temp$varexp, breaks=10)
o <- group_by(temp, vbin, cond_bonf_bonf2, nsig_disc_bonf) %>%
summarise(n=n())
o$what <- "simulation"

e <- tibble(cond_bonf_bonf2=5, nsig_disc_bonf=11, n=1, what="empirical", vbin = unique(temp$vbin))
oe <- bind_rows(o,e)
oe$rr <- oe$cond_bonf_bonf2 / oe$nsig_disc_bonf
oe$rr[is.na(oe$rr)] <- 0

ggplot(oe, aes(x=nsig_disc_bonf, y=rr)) +
geom_point(aes(color=what, size=n)) +
labs(x="Number discovered", y="Proportion replicated") +
facet_grid(. ~ vbin)
```
